---
bibliography: LTREB_jamaica.bib
csl: Styles/GlobalChangeBiology.csl
output:
  word_document:
    reference_docx: Styles/word-styles-reference-01.docx
  pdf_document: default
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    theme: paper
---

#Results
Figure 1. Yearly precipitation decreased significantly between 2007 and 2017. Dashed line indcates mean from 1980-2018. Study period highlighted with gray polygon. 

```{r figure 1, fig.height = 6, fig.width = 6, dpi = 300, echo = FALSE}
#rainfall <- read.csv("../Data/PrecipMonthly2007_2017.csv")
#par(bty = "l")
#plot(tapply(rainfall$MonthlyPrecip,rainfall$year,sum),
#     ylim = c(500,3000),
#     yaxt = "n", xaxt = "n",las = 1,
#     ylab = "Precipitation (mm)",
#     xlab = "Year", pch = 19, cex = 1.25)
#axis(1,at = 1:10,labels = 2007:2016)
#axis(2,las = 1)
MoBay <- read.csv("../Data/MontegoBay_WeatherStation.csv")
MoBay[MoBay == "NaN"] <- NA

MoBay$date <- as.POSIXct(MoBay$DATE, format = "%Y-%m-%d")

MoBay$Year <- format(MoBay$date,"%Y")
MoBay$Month <- format(MoBay$date, "%m")


MonthTemp <- tapply(MoBay$TAVG, list(MoBay$Year,MoBay$Month), FUN = mean, na.rm = TRUE)
MonthPrecip <- tapply(MoBay$PRCP, list(MoBay$Year,MoBay$Month), FUN = mean, na.rm = TRUE)

# Rolling coefficient of variation
#zoo:::rollapply(MonthPrecip[,3],width = 5, by = 5, FUN = function(x){sd(x,na.rm = #TRUE)/mean(x,na.rm = TRUE)*100}, align = "right")

YearlyTemp <- tapply(MoBay$TAVG, MoBay$Year, mean, na.rm = TRUE)
YearlyPrec <- tapply(MoBay$PRCP, MoBay$Year, sum, na.rm = TRUE)

mean30yrPPT <- mean(YearlyPrec[1:30])

mean30yrTEMP <- apply(MonthTemp[1:30,],2,mean,na.rm = TRUE)
mean30yrPRECIP <- apply(MonthPrecip[1:30,],2,median,na.rm = TRUE)

#par(mfrow = c(2,1))
par(mar = c(2,4,2,2))
plot(YearlyPrec[8:length(YearlyPrec)],type = "o",pch = 19, cex = 1.2,
     ylab = "Annual Precipitation (cm)",yaxt = "n",xaxt = "n",ylim = c(0,120))
axis(2, las = 2)
axis(1, las = 1, labels = c(1980:2018),at = c(1:39))
polygon(x = (c(36,45,45,36)-7), y = c(0.0,0.0,0.3,0.3), col = c("gray88"), border = "gray88")
points(MonthPrecip[8:nrow(MonthPrecip),3], type = "o", pch = 19, cex = 1.2)
abline(h = mean30yrPRECIP[3], lty = 2)
fit <- lm(MonthPrecip[rownames(MonthPrecip) %in% c(2008:2017),3]~c(36:45))
y0 <- 36*fit$coefficients[2]+fit$coefficients[1]
y1 <- 45*fit$coefficients[2]+fit$coefficients[1]
segments(x0 = (36-7), x1 = (45-7), y0 = y0 , y1 = y1, lwd = 3)


par(mfrow = c(2,1))
par(mar = c(2,4,2,2))
plot(MonthPrecip[8:nrow(MonthPrecip),3],ylim = c(0,0.3),type = "o",pch = 19, cex = 1.2,
     ylab = "March Precipitation (cm)",yaxt = "n",xaxt = "n")
axis(2, las = 2)
axis(1, las = 1, labels = c(1980:2018),at = c(1:39))
polygon(x = (c(36,45,45,36)-7), y = c(0.0,0.0,0.3,0.3), col = c("gray88"), border = "gray88")
points(MonthPrecip[8:nrow(MonthPrecip),3], type = "o", pch = 19, cex = 1.2)
abline(h = mean30yrPRECIP[3], lty = 2)
fit <- lm(MonthPrecip[rownames(MonthPrecip) %in% c(2008:2017),3]~c(36:45))
y0 <- 36*fit$coefficients[2]+fit$coefficients[1]
y1 <- 45*fit$coefficients[2]+fit$coefficients[1]
segments(x0 = (36-7), x1 = (45-7), y0 = y0 , y1 = y1, lwd = 3)

par(mar = c(2,4,2,2))
plot(MonthTemp[8:nrow(MonthTemp),3],ylim = c(76,83),type = "o",pch = 19, cex = 1.2,
     ylab = "March Temperature (F)",yaxt = "n",xaxt = "n")
axis(2, las = 2)
axis(1, las = 1, labels = c(1980:2018),at = c(1:39))
polygon(x = (c(36,45,45,36)-7), y = c(76,76,83,83), col = c("gray88"), border = "gray88")
points(MonthTemp[8:nrow(MonthTemp),3], type = "o", pch = 19, cex = 1.2)
abline(h = mean30yrTEMP[3], lty = 2)
fit <- lm(MonthTemp[rownames(MonthTemp) %in% c(2008:2017),3]~c(36:45))
y0 <- 36*fit$coefficients[2]+fit$coefficients[1]
y1 <- 45*fit$coefficients[2]+fit$coefficients[1]
segments(x0 = (36-7), x1 = (45-7), y0 = y0 , y1 = y1, lwd = 3)

```


```{r echo = FALSE}
# Yearly temp, precip and SOI 

MoBay <- read.csv("../Data/MontegoBay_WeatherStation.csv")

MoBay$date <- as.POSIXct(MoBay$DATE, format = "%Y-%m-%d")

MoBay$Year <- format(MoBay$date,"%Y")
MoBay$Month <- format(MoBay$date, "%m")


MonthTemp <- tapply(MoBay$TAVG, list(MoBay$Year,MoBay$Month), FUN = mean, na.rm = TRUE)
MonthPrecip <- tapply(MoBay$PRCP, list(MoBay$Year,MoBay$Month), FUN = mean, na.rm = TRUE)

SOI <- as.matrix(read.table("https://crudata.uea.ac.uk/cru/data/soi/soi.dat",header = FALSE))

colnames(SOI) <- c("Year","Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec","Annual")

SOI[SOI==-99.99]<-NA

SOI <- SOI[SOI[,1] > 2007,]

# Years 2008-2017 mean Jan-Mar
T<-apply(MonthTemp[11:20,1:3],1,mean)
P<-apply(MonthPrecip[11:20,1:3],1,sum,na.rm = TRUE)
S <- SOI[1:10,14] 

StandTemp <- (T-mean(T))/sd(T)
StandPrecip <- (P-mean(P))/sd(P)
SOIindex <- S

```

```{r,dpi = 300, echo = FALSE}
sppRDS <- list.files("../Results/",pattern = "*streamline.rds",full.names = TRUE)
sppSurv <- lapply(sppRDS,readRDS)
spp <- substr(sppRDS,12,15)
names(sppSurv)<-spp
```

Survival estimates for resident and migratory songbirds wintering in Jamaica. 

Each individual species is shown in gray. Mean survival of migrants (black points, dotted lines) and residents (triangles, dotted lines) are shown over the 10 year sampling period. Error bars represent the 95% credible interval. Survival estimates were generated using a spatially explicit mark-recapture model. 


```{r, echo = FALSE, dpi = 300}
mig <- c(1,0,1,1,1,0,0,0,0,0,0,1,1,1,0,1,1,0,1,0,1,0,0)
```

```{r, echo = FALSE, dpi = 300}
# AMRE
par(bty = "l")
plot(NA, xlim = c(1,10), ylim = c(0,1),yaxt = 'n',xaxt='n',
     ylab = expression(phi), main = "", xlab = "")
axis(2,las = 2)
axis(1, at = 1:10, labels = 2008:2017)
#legend(x=1,y=3,legend = c("Mangrove","Logwood"), pch = c(19,17),lty=c(2,1),bty = "n")
for(i in 1:length(spp)){
points(sppSurv[[i]]$mean$phi,
       pch = ifelse(mig[i]==1,19,17),
       lty = 2,
       col = ifelse(mig[i]==1,"gray88","gray88"),
       cex = ifelse(mig[i]==1,1,1.25),
       )
lines(sppSurv[[i]]$mean$phi,
      col = "gray88",
      lty = ifelse(mig[i] == 1, 1,2))
}
meanSurv <- lapply(sppSurv[which(mig == 1)], FUN = function(x){x$mean$phi})
migEsts <- apply(do.call(rbind,meanSurv),2,FUN = function(x){quantile(x,probs = c(0.5,0.025,0.975))})


meanRes <- lapply(sppSurv[which(mig == 0)], FUN = function(x){x$mean$phi})
resEsts <- apply(do.call(rbind,meanRes),2,FUN = function(x){quantile(x,probs = c(0.5,0.025,0.975))})

points(migEsts[1,], pch = 19, col = "black", cex = 1.25)
lines(migEsts[1,], lty = 2)
segments(x0 = 1:10, x1 = 1:10, y0 = migEsts[2,], y1 = migEsts[3,])

points(1.1:10.1,resEsts[1,], col = "black", cex = 1.25, pch = 17)
lines(1.1:10.1,resEsts[1,])
segments(x0 = 1.1:10.1, x1 = 1.1:10.1, y0 = resEsts[2,], y1 = resEsts[3,])
legend(1,0.2, legend = c("Resident","Migrant"), pch = c(17,19),lty=c(1,2),bty = "n",col = c("black","black"))
```

```{r, dpi = 300, echo = FALSE}
# read in data
data.files <- list.files("../Data/", pattern = "windata", full.names = TRUE)
win.data <- lapply(data.files, readRDS)

data.spp <- substr(data.files,9,12)
```


```{r, echo = FALSE,dpi = 300}
removespp <- spp[c(5,6,8,10,12,15,18)]
```

*The next series of figures uses only species with 30 or more total recaptures over the 10 year period.*

It includes the following migrants:
```{r,echo = FALSE}
spp[!(spp %in% removespp) & mig == 1]
```

It includes the following residents:
```{r,echo = FALSE}
spp[!(spp %in% removespp) & mig == 0]
```


```{r, echo = FALSE}
mig <- mig[!(spp %in% removespp)]

spp1 <- spp[!(spp %in% removespp)]

win.data <- win.data[match(spp,data.spp)]
```


Below is the relationship between survival and the Annual Southern Oscillation Index. Positive values are associated with La Nina, while negative are associated with El Nino. 

```{r, echo = FALSE, dpi = 300}
# sum(apply(win.data[[]]$Ycat, 3, FUN = function(x){sum(x != 81)}))

sPredict <- array(NA,c(100,length(spp1)))
plot(NA, ylim = c(0,1), xlim = c(-1.5,1.5), bty= "l", las = 1, xlab = "SOI index", ylab = expression(phi))
for(i in 1:length(spp1)){
points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$soi, pch = ifelse(mig[i] == 1, 19,17), col = "gray88")
stdMetric <- seq(min(win.data[[1]]$soi),max(win.data[[1]]$soi),,100)
lphi <- log(sppSurv[[i]]$mean$phi/(1-sppSurv[[i]]$mean$phi))

sPredict[,i] <- (1 / (1+exp(-((mean(lphi)+sppSurv[[i]]$mean$beta.phi[1]*0+ # precip[y]+
                                  sppSurv[[i]]$mean$beta.phi[2]*0+ #temp[y]+
                                  sppSurv[[i]]$mean$beta.phi[3]*0+ #(precip[y]*temp[y])+
                                  sppSurv[[i]]$mean$beta.phi[4]*stdMetric)))))
points(sPredict[,i] ~ stdMetric, type = "l", lty = 1-mig[i], lwd = 0.25, col = "gray")    
# points(sPredict[,i] ~ stdMetric, type = "l", lty = 2-mig[i])           
# points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$soi, ylim = c(0,1), pch = 19, col = 2-mig[i])
}

migSOI <- apply(sPredict[,which(mig == 1)],1,mean)
migError <- apply(sPredict[,which(mig == 1)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})
resSOI <- apply(sPredict[,which(mig == 0)],1,mean)
resError <- apply(sPredict[,which(mig == 0)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})

polygon(x = c(stdMetric,rev(stdMetric)), y = c(migError[1,],rev(migError[2,])), border = "gray", col = rgb(220,220,220,100,max = 255))
polygon(x = c(stdMetric,rev(stdMetric)), y = c(resError[1,],rev(resError[2,])), border = rgb(180,180,180,100,max = 255), col = rgb(200,200,200,100,max = 255))
points(migSOI ~ stdMetric, type = "l", lty = 1, lwd = 2) 
points(resSOI ~ stdMetric, type = "l", lty = 2, lwd = 2) 

legend("bottomleft", legend = c("Migrants","Residents"), lty = c(1,2), fill = c(rgb(220,220,220,100,m = 255),rgb(200,200,200,200,m = 255)), bty = "n")
```



Below is the relationship in annual survival and mean temperature between January - March for the years 2008-2017.


```{r, echo = FALSE, dpi = 300}
stdMetric <- seq(min(win.data[[1]]$temp),max(win.data[[1]]$temp),,100)
sPredict <- array(NA,c(100,23))
plot(NA, ylim = c(0,1), xlim = c(min(stdMetric),max(stdMetric)), bty= "l", las = 1, xlab = expression(paste("Temperature [",degree,"C]")), ylab = expression(phi),
     xaxt = "n")
axis(1, at = seq(-2.5,1,0.5), labels = round(((seq(-2.5,1.0,0.5)*sd(T)+mean(T))-32)*(5/9),2))

for(i in 1:length(spp1)){

points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$temp, pch = ifelse(mig[i] == 1, 19,17), col = "gray88")
  
lphi <- log(sppSurv[[i]]$mean$phi/(1-sppSurv[[i]]$mean$phi))

sPredict[,i] <- (1 / (1+exp(-((mean(lphi)+sppSurv[[i]]$mean$beta.phi[1]*0+ # precip[y]+
                                  sppSurv[[i]]$mean$beta.phi[2]*stdMetric+ #temp[y]+
                                  sppSurv[[i]]$mean$beta.phi[3]*0+ #(precip[y]*temp[y])+
                                  sppSurv[[i]]$mean$beta.phi[4]*0)))))

points(sPredict[,i] ~ stdMetric, type = "l", lty = 1-mig[i], lwd = 0.25, col = "gray")    
}

migSOI <- apply(sPredict[,which(mig == 1)],1,mean)
migError <- apply(sPredict[,which(mig == 1)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})
resSOI <- apply(sPredict[,which(mig == 0)],1,mean)
resError <- apply(sPredict[,which(mig == 0)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})

polygon(x = c(stdMetric,rev(stdMetric)), y = c(migError[1,],rev(migError[2,])), border = "gray", col = rgb(220,220,220,100,max = 255))
polygon(x = c(stdMetric,rev(stdMetric)), y = c(resError[1,],rev(resError[2,])), border = rgb(180,180,180,100,max = 255), col = rgb(200,200,200,100,max = 255))
points(migSOI ~ stdMetric, type = "l", lty = 1, lwd = 2) 
points(resSOI ~ stdMetric, type = "l", lty = 2, lwd = 2) 

legend("bottomright", legend = c("Migrants","Residents"), lty = c(1,2), fill = c(rgb(220,220,220,100,m = 255),rgb(200,200,200,200,m = 255)), bty = "n")

```


Below is the relationship between Jan - Mar precipitation (cm) and survival. 


```{r, echo = FALSE, dpi = 300}
stdMetric <- seq(min(win.data[[1]]$precip),max(win.data[[1]]$precip),,100)
sPredict <- array(NA,c(100,length(spp1)))
plot(NA, ylim = c(0,1), xlim = c(min(stdMetric),max(stdMetric)), bty= "l", las = 1, xlab = "Precipitation (cm)", ylab = expression(phi), xaxt = "n")
axis(1, at = seq(-1.5,2.5,0.5), labels = round(seq(-1.5,2.5,0.5)*sd(P)+mean(P),2))

for(i in 1:length(spp1)){

points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$precip, pch = ifelse(mig[i] == 1, 19,17), col = "gray88")
  
lphi <- log(sppSurv[[i]]$mean$phi/(1-sppSurv[[i]]$mean$phi))

sPredict[,i] <- (1 / (1+exp(-((mean(lphi)+sppSurv[[i]]$mean$beta.phi[1]*stdMetric+ # precip[y]+
                                  sppSurv[[i]]$mean$beta.phi[2]*0+ #temp[y]+
                                  sppSurv[[i]]$mean$beta.phi[3]*0+ #(precip[y]*temp[y])+
                                  sppSurv[[i]]$mean$beta.phi[4]*0)))))

points(sPredict[,i] ~ stdMetric, type = "l", lty = 1-mig[i], lwd = 0.25, col = "gray")           
# points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$soi, ylim = c(0,1), pch = 19, col = 2-mig[i])
}

migSOI <- apply(sPredict[,which(mig == 1)],1,mean)
migError <- apply(sPredict[,which(mig == 1)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})
resSOI <- apply(sPredict[,which(mig == 0)],1,mean)
resError <- apply(sPredict[,which(mig == 0)],1, FUN = function(x){quantile(x,probs = c(0.025, 0.975))})

polygon(x = c(stdMetric,rev(stdMetric)), y = c(migError[1,],rev(migError[2,])), border = "gray", col = rgb(220,220,220,100,max = 255))
polygon(x = c(stdMetric,rev(stdMetric)), y = c(resError[1,],rev(resError[2,])), border = rgb(180,180,180,100,max = 255), col = rgb(200,200,200,100,max = 255))
points(migSOI ~ stdMetric, type = "l", lty = 1, lwd = 2) 
points(resSOI ~ stdMetric, type = "l", lty = 2, lwd = 2) 

legend("bottomright", legend = c("Migrants","Residents"), lty = c(1,2), fill = c(rgb(220,220,220,100,m = 255),rgb(200,200,200,200,m = 255)), bty = "n")

```


The following plots shows the results for each species. 


```{r, echo = FALSE, dpi = 300}
stdMetric <- seq(min(win.data[[1]]$precip),max(win.data[[1]]$precip),,20)

sPredict <- array(NA,c(20,3000,length(spp1)))

for(i in 1:length(spp1)){
  
lphi <- log(sppSurv[[i]]$sims.list$phi/(1-sppSurv[[i]]$sims.list$phi))
lphi <- apply(lphi,1,mean)
sPredict[,,i] <- (1 / (1+exp(-((lphi+sppSurv[[i]]$sims.list$beta.phi[,1]*stdMetric+ # precip[y]+
                                  sppSurv[[i]]$sims.list$beta.phi[,2]*0+ #temp[y]+
                                  sppSurv[[i]]$sims.list$beta.phi[,3]*0+ #(precip[y]*temp[y])+
                                  sppSurv[[i]]$sims.list$beta.phi[,4]*0)))))

#points(sPredict[,i] ~ stdMetric, type = "l", lty = 1-mig[i], lwd = 0.25, col = "gray")           
# points(sppSurv[[i]]$mean$phi ~ win.data[[1]]$soi, ylim = c(0,1), pch = 19, col = 2-mig[i])
}
meanSurv_precip <- apply(sPredict,c(1,3),FUN = function(x){quantile(x,probs = c(0.5,0.025,0.975))})

for(i in 1:length(spp1)){
plot(NA, ylim = c(0,1), xlim = c(min(stdMetric),max(stdMetric)), bty= "l", las = 1, xlab = "Precipitation (cm)", ylab = expression(phi), xaxt = "n",
     main = spp1[i])
axis(1, at = seq(-1.5,2.5,0.5), labels = round(seq(-1.5,2.5,0.5)*sd(P)+mean(P),2))
polygon(x = c(stdMetric, rev(stdMetric)), y = c(meanSurv_precip[2,,i], rev(meanSurv_precip[3,,i])), col = "gray", border = "gray")
points(x = stdMetric, y = meanSurv_precip[1,,i], type = "l", lwd = 2)
}

```

Below are a ton of plots that show yearly survival estimates, recruitment, density on the plots, etc. 


```{r, echo = FALSE, dpi = 300}
for(i in 1:length(spp)){
par(bty = "l")
plot(NA, xlim = c(1,10), ylim = c(0,1),yaxt = 'n',xaxt='n',
     ylab = expression(phi), main = paste(spp[i],"surival"), xlab = "")
axis(2,las = 2)
axis(1, at = 1:10, labels = 2008:2017)
legend(x=1,y=3,legend = c("Mangrove","Logwood"), pch = c(19,17),lty=c(2,1),bty = "n")
points(sppSurv[[i]]$mean$phi,pch = ifelse(mig[i]==1,19,17),type = "o",lty = 2,col = ifelse(mig[i]==1,"black","gray"), cex = ifelse(mig[i]==1,1.25,2))

par(bty = "l")
plot(NA, ylim = c(min(sppSurv[[i]]$q2.5$N,na.rm = TRUE),max(sppSurv[[i]]$q97.5$N,na.rm = TRUE)), xlim = c(1,10),yaxt = 'n',xaxt='n',
     ylab = "Number individuals", main = paste(spp[i],"Population"), xlab = "")
axis(2,las = 2)
axis(1, at = 1:10, labels = 2008:2017)
points(sppSurv[[i]]$mean$N)
segments(x0 = 1:10,
         x1= 1:10,
         y0 = sppSurv[[i]]$q2.5$N,
         y1 = sppSurv[[i]]$q97.5$N)

par(bty = "l")
plot(NA, ylim = c(min(sppSurv[[i]]$q2.5$R,na.rm = TRUE),max(sppSurv[[i]]$q97.5$R,na.rm = TRUE)), xlim = c(1,10),yaxt = 'n',xaxt='n',
     ylab = "Number recruits", main = paste(spp[i],"Recruitment"), xlab = "")
axis(2,las = 2)
axis(1, at = 1:10, labels = 2008:2017)
points(sppSurv[[i]]$mean$R)
segments(x0 = 1:10,
         x1= 1:10,
         y0 = sppSurv[[i]]$q2.5$R,
         y1 = sppSurv[[i]]$q97.5$R)

par(bty = "l")
plot(NA, xlim = c(1,10), ylim = c(0,3),yaxt = 'n',xaxt='n',
     ylab = "Density (N/ha)", main = spp[i], xlab = "")
axis(2,las = 2)
axis(1, at = 1:10, labels = 2008:2017)
legend(x=1,y=3,legend = c("Mangrove","Logwood"), pch = c(19,17),lty=c(2,1),bty = "n")
points(sppSurv[[i]]$mean$Dplots[1,],pch = 19, type = "o",lty = 2)
points(sppSurv[[i]]$mean$Dplots[2,],pch = 17, type = "o")
points(sppSurv[[i]]$mean$Dplots[3,],pch = 19, type = "o",lty = 2)
points(sppSurv[[i]]$mean$Dplots[4,],pch = 17, type = "o")
points(sppSurv[[i]]$mean$Dplots[5,],pch = 19, type = "o",lty = 2)
}

```

Just for fun I threw in some cool additional data - activity centers. 

```{r, echo = FALSE, dpi = 300}

###################################################################################
# Read in shapefiles of the plots #
library(raster)

JAMplots <- raster::shapefile("../Spatial_Layers/JAMplots_all.shp")

# identify plots as shapefiles #
MM <- JAMplots[JAMplots$PLOT == "MM",]
LI <- JAMplots[JAMplots$PLOT == "LI",]
MI <- JAMplots[JAMplots$PLOT == "MI",]
LII <- JAMplots[JAMplots$PLOT == "LII",]
MII <- JAMplots[JAMplots$PLOT == "MII",]

# create a single list of the plots #
plots <- list(MM,LI,MI,LII,MII)

# generate convex hull around each plot 
plotpolys <- lapply(plots,rgeos::gConvexHull)

plotspecificarea <- lapply(plotpolys,FUN = function(x){rgeos::gArea(x)/10000})

# 5 x 5 m raster of plot extent
JAMraster <- raster(extent(JAMplots),res=c(5,5))

# set values to 0 - i.e., not sampled 
JAMraster[]<-0

# make rasters for each plot and give them values 1 - 5
plotrasters <- vector("list",5)
for(i in 1:5){
plotrasters[[i]] <- rasterize(plotpolys[[i]],JAMraster)
plotrasters[[i]][plotrasters[[i]]==1]<-i
}

# Calculate area of the 5 plots in hectares
(area(plotpolys[[1]])+area(plotpolys[[2]])+area(plotpolys[[3]])+area(plotpolys[[4]])+area(plotpolys[[5]]))/10000

# habmatrix 
study.area <- sum(stack(c(JAMraster,plotrasters)),na.rm = TRUE)

# create matrix
habmat <- as.matrix(study.area)

# Flip the plot raster to match the net locations 
study.flip <- flip(raster(habmat,xmn=0, xmx=ncol(habmat), ymn=0, ymx=nrow(habmat)),"y")
```

Estimated activity center for 3 American Redstarts using the recapture information in LTREB. The colors represent the different plots. The color ramps represent the posterior estimate for individual activity centers. 

```{r, echo = FALSE, dpi = 300}
study.flip[study.flip == 0]<-NA
par(bty = "n", mar = c(0,0,0,0))
plot(study.flip, legend = FALSE, axes = FALSE, col = rev(bpy.colors(5)))
xcoords <- sppSurv$AMRE$sims.list$S[,3,2,1]*sppSurv$AMRE$sims.list$z[,3,1]
ycoords <- sppSurv$AMRE$sims.list$S[,3,1,1]*sppSurv$AMRE$sims.list$z[,3,1] 
xcoords <- xcoords[xcoords>0]
ycoords <- ycoords[ycoords>0]
# points(ycoords~xcoords, pch = ".")
AC <- raster(ks::kde(cbind(xcoords,ycoords)))
AC[AC<=0] <- NA
plot(AC,add = TRUE, alpha = 0.5, legend = FALSE)

xcoords <- sppSurv$AMRE$sims.list$S[,50,2,1]*sppSurv$AMRE$sims.list$z[,50,1]
ycoords <- sppSurv$AMRE$sims.list$S[,50,1,1]*sppSurv$AMRE$sims.list$z[,50,1] 
xcoords <- xcoords[xcoords>0]
ycoords <- ycoords[ycoords>0]
# points(ycoords~xcoords, pch = ".")
AC <- raster(ks::kde(cbind(xcoords,ycoords)))
AC[AC<=0] <- NA
plot(AC,add = TRUE, alpha = 0.5, legend = FALSE)
```


